Iroko: Predictive Datacenter Congestion Control

### 2017-12-12 Documents/Report/intro.tex
Edits

Developments in the past decade have changed the general networking 
environment. Data centers have emerged as an exciting new research frontier, 
posing novel design challenges and opportunities. Driven by minimization of 
costs and maximization of compute power, they must run at the highest 
possible utilization to achieve the ideal compute/cost ratio.

@@

Despite the innovation potential of data centers, TCP has dominated as 
the congestion control algorithm of choice.
Still, TCP is a \textit{reactive protocol}, only responding to signals of 
congestion. A substantial and noticeable increase of loss and latency in the 
network already portends a problem. Overflowing queues in forwarding elements 
or mismatched hardware capabilities imply that traffic has not been optimally 
managed.
Ideally, a network should always be "zero-queue", i.e., latency 
will merely be induced by propagation, and not queuing delay.

TCP's algorithm is a best-effort solution to globally maximize utilization 
under the fairness principle. This design has not fundamentally changed 
over the last thirty years, and new research advancements raise the question if TCP can still be considered to be the most sensible option.~\cite{pcc,bbr,perc}

@@

In parallel, admission-control-based approaches have gained traction as a 
viable alternative to the conventional burst-and-backoff mechanism of 
TCP.~\cite{expresspass, fastpass, perc}
While idea of admission control and service guarantees in networks is not 
new~\cite{access_limit, access_limit2}, these designs traditionally 
aim to assure quality and bandwidth guarantees in a contentious, decentralized, 
and untrusted environments such as the internet. In a data center, these 
conditions do not apply. End-hosts are generally considered reliable and 
restricted in behaviour, which allows for great simplification of enforcement 
and prioritization policies. Admission-based systems can give administrators 
the advantage of tight control over the data center, mitigating the effects of 
queue-buildup and bursts.

Although it is unclear how predictable data center traffic truly is, various 
works have shown that repeated patterns may 
exist.\cite{msr_dc,fb_dc,traffic,microte}
If data center traffic exhibits predictable patterns, it may be 
possible to devise a proactive mechanism which forecasts the traffic matrix of 
next iterations, and adjusts traffic accordingly.

@@

This project report details our exploration of such a centralized, 
proactive flow scheduler. We asked ourselves the following research questions:
\begin{enumerate}
    \item What are the requirements for such a centralized, predictive 
    scheduler to succeed?
    \item Is it possible to preemptively regulate a network by analyzing global 
    traffic patterns?
    \item What latency, packet loss and utilization are we able to achieve?
\end{enumerate}
In the scope of this report, we outline the design and implementation of our 
flow scheduler and learning approach. We perform an analysis of its 
effectiveness compared to DCTCP and Hedera and conclude with a discussion of 
the benefits and drawbacks of centralized learning and admission control in the 
data center context.

>>>>

Developments in the past decade have changed the general networking 
environment. Datacenters have emerged as an exciting new research frontier, 
posing novel design challenges and opportunities. Driven by minimization of 
costs and maximization of compute power, they must run at the highest 
possible utilization to achieve the ideal compute/cost ratio.

@@

Despite the innovation potential of datacenters, TCP has dominated as 
the congestion control algorithm of choice.
Still, TCP is a \textit{reactive protocol}, only responding to signals of 
congestion. A substantial and noticeable increase of loss and latency in the 
network already portends a problem. Overflowing queues in forwarding elements 
or mismatched hardware capabilities imply that traffic has not been optimally 
managed.
Ideally, a network should always be "zero-queue", i.e., latency 
will merely be induced by propagation, and not queuing delay.

TCP's algorithm is a best-effort solution to globally maximize utilization 
under the fairness principle. This design has not fundamentally changed 
over the last thirty years, and new research advancements raise the question whether TCP
can still be considered the most sensible option.~\cite{pcc,bbr,perc}

@@

In parallel, admission-control-based approaches have gained traction as a 
viable alternative to the conventional burst-and-backoff mechanism of 
TCP.~\cite{expresspass, fastpass, perc}
While the idea of admission control and service guarantees in networks is not 
new~\cite{access_limit, access_limit2}, these designs traditionally 
aim to assure quality and bandwidth guarantees in contentious, decentralized, 
and untrusted environments such as the internet. In a datacenter, these 
conditions do not apply. End-hosts are generally considered reliable and 
restricted in behaviour, which allows for great simplification of enforcement 
and prioritization policies. Admission-based systems can give administrators 
the advantage of tight control over the datacenter, mitigating the effects of 
queue-buildup and bursts.

Although it is unclear how predictable datacenter traffic truly is, various 
works have shown that repeated patterns may 
exist.\cite{msr_dc,fb_dc,traffic,microte}
If datacenter traffic exhibits predictable patterns, it may be 
possible to devise a proactive mechanism which forecasts the traffic matrix of 
next iterations, and adjusts traffic accordingly.

@@

This project report details our exploration of such a centralized, 
proactive flow scheduler. We asked ourselves the following research questions:
\begin{enumerate}
    \item What are the requirements for such a centralized, predictive 
    scheduler to succeed?
    \item Is it possible to preemptively regulate a network by analyzing global 
    traffic patterns?
    \item What latency, packet loss and utilization are we able to achieve?
\end{enumerate}
In the scope of this report, we outline the design and implementation of our 
flow scheduler and learning approach. We perform an analysis of its 
effectiveness compared to DCTCP and Hedera and conclude with a discussion of 
the benefits and drawbacks of centralized learning and admission control in the 
datacenter context.

### 2017-12-12 Documents/Report/related_work.tex
Edits

In this project, we compare Iroko to two particular TCP protocols of interest.  
The first, Data Center TCP (DCTCP)~\cite{dctcp}, is a congestion control 
algorithm developed to reduce bufferbloat in data center networks. DCTCP 
leverages Explicit Congestion Notifications (ECNs) as early warning signal to 
identify queue buildup and adjusts its sending rate to the amount of ECN-marked 
packets on the receiving side. As part of the static TCP congestion control 
family, DCTCP is fully decentralized, does not limit host sending rates in 
advance and relies on a static congestion control model.

The second algorithm we compare to is TCP CUBIC~\cite{cubic}, the default 
congestion control protocol of the Linux kernel. CUBIC uses a cubic function   
to increase window size and serves as the baseline for Iroko. It is intended to 
work well over high bandwidth, high latency networks.

@@

MicroTE~\cite{microte} is a central scheduler working under the 
assumption that a subset of short-term traffic is predictable. The controller 
isolates flows which it determined to be predictable and install a custom 
optimal routing entry. The remaining unpredictable traffic is routed according 
to ECMP striping.

@@

FastPass~\cite{fastpass} employs an arbiter mechanism, in which hosts request 
access to the network resources. For any new flow, the FastPass shim layer will 
send a "admission request" to a central controller, which chooses an optimal 
allocated time slot and path. Once an applications has been assigned this path, 
it is guaranteed to experience zero queuing latency and jitter during 
transmission. This "zero-loss" network provides application with a stable and 
reliable transmission experience. However, FastPass still works reactively and 
only operates in incoming requests. Queueing is thus pushed to end-hosts and 
the arbiter itself.

@@

Iroko builds heavily on the contributions of both FastPass and ExpressPass. 
However, we intend to go further and schedule rate-limits ahead of time in 
order to reduce sending delays and host starvation of admission-control systems 
substantially.

\subsection{Predicting and Learning Network Traffic}
A new direction of networking research is study on the predictability and 
trainability of network traffic.~\cite{throwdown, learning_tcp}
A common assumption is that data center traffic is largely unpredictable and 
untrainable.~\cite{fb_dc}
However, Winstein and Balakrishnan,~2013\cite{remy} demonstrated that it is 
possible to learn near-optimal TCP congestion control parameters for various 
network topologies. The system, Remy, represents an offline learning algorithm 
inferring optimal TCP parameters based on pre-supplied topology and 
traffic-model.
While Remy only performed static offline training, PCC~\cite{pcc,throwdown} 
aims to develop a dynamic and actively improving congestion control algorithm. 
PCC makes no prior assumption about the network and continuously reiterates over
its host-local optimization function to find optimal parameters.

As part of the contribution of MicroTE, Benson et 
al.,~2014~\cite{microte} analyzed data center traffic and 
inferred that around 30\% of short-lived traffic is predictable over a period 
of 1-2 seconds.

>>>>

In this project, we compare Iroko to two particular TCP protocols of interest.  
The first, Data Center TCP (DCTCP)~\cite{dctcp}, is a congestion control 
algorithm developed to reduce bufferbloat in datacenter networks. DCTCP 
leverages Explicit Congestion Notifications (ECNs) as early warning signals to 
identify queue buildup and adjusts its sending rate to the amount of ECN-marked 
packets on the receiving side. As part of the static TCP congestion control 
family, DCTCP is fully decentralized, does not limit host sending rates in 
advance and relies on a static congestion control model.

The second algorithm we compare is TCP CUBIC~\cite{cubic}, the default 
congestion control protocol of the Linux kernel. CUBIC uses a cubic function   
to increase window size and serves as the baseline for Iroko. It is intended to 
work well over high bandwidth, high latency networks.

@@

MicroTE~\cite{microte} is a central scheduler working under the 
assumption that a subset of short-term traffic is predictable. The controller 
isolates flows which it determined to be predictable and installs a custom 
optimal routing entry. The remaining unpredictable traffic is routed according 
to ECMP striping.

@@

FastPass~\cite{fastpass} employs an arbiter mechanism, in which hosts request 
access to the network resources. For any new flow, the FastPass shim layer will 
send a "admission request" to a central controller, which chooses an optimal 
allocated time slot and path. Once an applications has been assigned this path, 
it is guaranteed to experience zero queuing latency and jitter during 
transmission. This "zero-loss" network provides applications with a stable and 
reliable transmission experience. However, FastPass still works reactively and 
only operates on incoming requests. Queueing is thus pushed to end-hosts and 
the arbiter itself.

@@

Iroko builds heavily on the contributions of both FastPass and ExpressPass. 
However, we intend to go further and schedule rate-limits ahead of time in 
order to substantially reduce sending delays and host starvation of admission-control systems.

\subsection{Predicting and Learning Network Traffic}
A new direction of networking research is study on the predictability and 
trainability of network traffic.~\cite{throwdown, learning_tcp}
A common assumption is that datacenter traffic is largely unpredictable and 
untrainable.~\cite{fb_dc}
However, Winstein and Balakrishnan,~2013 \cite{remy} demonstrated that it is 
possible to learn near-optimal TCP congestion control parameters for various 
network topologies. The system, Remy, represents an offline learning algorithm 
inferring optimal TCP parameters based on pre-supplied topologies and 
traffic-models.
While Remy only performed static offline training, PCC~\cite{pcc,throwdown} 
aims to develop a dynamic and actively improving congestion control algorithm. 
PCC makes no prior assumption about the network and continuously reiterates over
its host-local optimization function to find optimal parameters.

As part of the contribution of MicroTE, Benson et 
al.,~2014~\cite{microte} analyzed datacenter traffic and 
inferred that around 30\% of short-lived traffic is predictable over a period 
of 1-2 seconds.
